---
title: "Assignment 4"
output: html_notebook
---
akankshamr@iisc.ac.in
                                                       MG 222: Assignment IV
                                                      Due Date: June 18, 2020
                                                      


The text file in http://www.mgmt.iisc.ac.in/CM/MG222/Data Files/gdp.data contains time series of Quarterly GDP of India from 1996-97:Q1, to 2013-14:Q2, in 2004-05 prices, split across eight sectors (provided by RBI, and as agreed upon by various other government bodies such as CSO, NSSO, Finance Ministry etc.) as well as the net GDP, in its last nine columns. The column names are self-explanatory and are thus not repeated here. In this class, there are five groups for Assignments purpose. Each one of these groups must choose only one of the following columns for this assignment (analysis of any other column of data will not be accepted), and no two groups must analyse the same time series. The acceptable

columns (as they appear in gdp.data) are as follows.  

Agriculture
• Manufacturing
• Trade, Hotels, Transport & Communication
• Community, Social & Personal Services
• Gross Domestic Product

                                          Community, Social & Personal Services


First leave aside the last four observations of 2012-13:Q3, 2012-13:Q4, 2013-14:Q1 and 2013-14:Q2, while you build the model. These will be used for checking out the quality of the forecasts based on the models you are about to build, and thus must not be used for modelbuilding purpose. That is, your models must be built based on the 66 observations from 1996-97:Q1 to 2012-13:Q2. Answer the following.


1. Build a regression model that you deem to be most appropriate for modeling the trend and seasonality present in your chosen time series. Do the residulas of this regression model satisfy all the required assumptions?                                                         [7+3=10]

```{r}
library("zoo")
library(dplyr)
library(car)
library(corrplot)
library(MASS)
library(lmtest) 
library("tseries")
library("forecast")
```

66 observations from 1996-97:Q1 to 2012-13:Q2

Year	Quarter	Agriculture	Mining	Manufacturing	Electricity,Gas&WaterSupply	Construction	Trade,Hotels,Transport&Communication
Finance,Insurance,RealEstate&BusinessServices	Community,Social&PersonalServices	GrossDomesticProduct

```{r}
data <- read.table("C:/Users/Akanksha/Documents/R Files/MG222/assignment4data.csv", header = T)  ## Read 
data
ts1<-ts(data$Community.Social.PersonalServices)
ts2<-ts1[1:66]
tsc=ts(ts2,deltat = 1/4)
tt=1:70
t=tt[1:66]

plot(t,tsc,type="n",main="Quarterly GDP of India from 1996-97 to 2012-13",xlab="Time",ylab="Quarterly GDP in Crores of Rs")
lines(t,tsc)
points(t,tsc,pch=20)
legend("topleft", c(" Community,Social&PersonalServices(Observed)","Trend"), lty=c(1,1), col=c('black','blue'), bty='n', cex=1)
decomposed.tsc<-decompose(tsc)
lines(t,decomposed.tsc$trend,col='blue',lwd=1.25)
plot(decomposed.tsc)
tsc
```
```{r}

```
1. Build a regression model that you deem to be most appropriate for modeling the trend and seasonality present in your chosen time series. Do the residulas of this regression model satisfy all the required assumptions? 

Model 1.  First Regression Model, which is simple linear function of t
```{r}
model1<-lm(tsc~t)
summary(model1)
par(mfrow = c(1, 2))
plot(model1,which=1:2)
res1<-residuals(model1)/sqrt(327.5)
boxplot(res1)

plot(t,res1)
lines(lowess(t,res1),col=2)

library(nortest)
ad.test(sres1)
cvm.test(sres1)
lillie.test(sres1)
pearson.test(sres1)
sf.test(sres1)
normtest(res1)
```
Model 2: Second Regression Model, which is a quadratic function of t and a linear t
```{r}
t1=t*t
model2<-lm(tsc~t+t1)
summary(model2)
plot(model2,which=1:2)
sres2<-residuals(model2)/sqrt(174.4)
boxplot(sres2)
plot(t,sres2)
lines(lowess(t,sres2),col=2)
plot(t1,sres2)
lines(lowess(t1,sres2),col=2)
library(nortest)
ad.test(sres2)
cvm.test(sres2)
lillie.test(sres2)
pearson.test(sres2)
sf.test(sres2)
normtest(sres2)
```
Model 3: Third Regression model,
```{r}
model3=lm(log(tsc)~t)
summary(model3)
plot(model3)
sres3<-residuals(model3)/sqrt(0.1337)
boxplot(sres3)
plot(t,sres3)
lines(lowess(t,sres3),col=2)
library(nortest)
ad.test(sres2)
cvm.test(sres2)
lillie.test(sres2)
pearson.test(sres2)
sf.test(sres2)
```
```{r}
model3=lm(log(tsc)~t+t1)
summary(model3)
plot(model3)
sres3<-residuals(model3)/sqrt(0.1337)
boxplot(sres3)
plot(t,sres3)
lines(lowess(t,sres3),col=2)
library(nortest)
ad.test(sres2)
cvm.test(sres2)
lillie.test(sres2)
pearson.test(sres2)
sf.test(sres2)
normtest(sres2)
```


Model 4: 
```{r}
w1=pi;
t1=t*t
model4=lm(tsc~t+t1+cos(w1*t)+sin(w1*0.5*t)+t1*sin(w1*0.5*t)+t1*cos(w1*0.5*t))
summary(model4)
plot(model4,which=1:2)
sres4=residuals(model4)/sqrt(138.9)
boxplot(sres4)
library(nortest)
ad.test(sres4)
cvm.test(sres4)
lillie.test(sres4)
pearson.test(sres4)
sf.test(sres4)
library(tseries)
res4=residuals(model4)
Box.test(res4)
res4=residuals(model4)
turning_point.test<-function(x)
{
     n<-length(x)
    turns<-0
    for(i in 2:(n-1))
    if((x[i-1]<x[i] && x[i]>x[i+1]) || (x[i-1]>x[i] && x[i]<x[i+1]))
    turns<-turns+1
    z<-(turns-2*(n-2)/3)/sqrt((16*n-29)/90)
    return(list(nturns=turns,z=z,p=2*(1-pnorm(abs(z)))))
     }
turning_point.test(res4)

diff_sign.test<-function(x)
{
n<-length(x)
y<-diff(x)
t<-length(y[y>0])
z<-(t-(n-1)/2)/sqrt((n+1)/12)
return(list(t=t,z=z,p=2*(1-pnorm(abs(z)))))
}
diff_sign.test(res4)

linear_trend.test<-function(x)
{
n<-length(x)
t<-0
for(i in 1:(n-1))
for(j in (i+1):n)
if(x[j]>x[i]) t<-t+1
z<-(t-n*(n-1)/4)/sqrt((n*(n-1)*(2*n+5)/72))
return(list(t=t,z=z,p=2*(1-pnorm(abs(z)))))
}
linear_trend.test(res4)

```
Model 5:
```{r}
w1=pi;
t1=t*t
model5=lm(tsc~t+t1+cos(w1*t)+cos(w1*0.5*t))
summary(model5)

model4<-model5
plot(model4,which=1:2)
sres4=residuals(model4)/sqrt(143)
boxplot(sres4)
library(nortest)
ad.test(sres4)
cvm.test(sres4)
lillie.test(sres4)
pearson.test(sres4)
sf.test(sres4)
library(tseries)
res4=residuals(model4)
Box.test(res4)
res4=residuals(model4)
turning_point.test<-function(x)
{
     n<-length(x)
    turns<-0
    for(i in 2:(n-1))
    if((x[i-1]<x[i] && x[i]>x[i+1]) || (x[i-1]>x[i] && x[i]<x[i+1]))
    turns<-turns+1
    z<-(turns-2*(n-2)/3)/sqrt((16*n-29)/90)
    return(list(nturns=turns,z=z,p=2*(1-pnorm(abs(z)))))
     }
turning_point.test(res4)

diff_sign.test<-function(x)
{
n<-length(x)
y<-diff(x)
t<-length(y[y>0])
z<-(t-(n-1)/2)/sqrt((n+1)/12)
return(list(t=t,z=z,p=2*(1-pnorm(abs(z)))))
}
diff_sign.test(res4)

linear_trend.test<-function(x)
{
n<-length(x)
t<-0
for(i in 1:(n-1))
for(j in (i+1):n)
if(x[j]>x[i]) t<-t+1
z<-(t-n*(n-1)/4)/sqrt((n*(n-1)*(2*n+5)/72))
return(list(t=t,z=z,p=2*(1-pnorm(abs(z)))))
}
linear_trend.test(res4)
```
```{r}
t=1:66
t1=t*t
w1=pi
reg_model=lm(tsc~t+t1+cos(w1*t)+cos(w1*0.5*t))
summary(reg_model)
plot(model4,which=1:2)
res=residuals(reg_model)/sqrt(143)  # standard residual
boxplot(res)
normtest<-function(x)
{
  library(nortest)
  s<-shapiro.test(x)
  ad<-ad.test(x)
  cvm<-cvm.test(x)
  ll<-lillie.test(x)
  sf<-sf.test(x)
  df<-data.frame(Method=c(s$method, ad$method, cvm$method, ll$method, sf$method),
                 P.Value=c(s$p.value, ad$p.value, cvm$p.value, ll$p.value, sf$p.value))
  df
}
normtest(res)
Box.test(res,lag=20)
Box.test(res,type="L",lag=20)
turning_point.test(res)
diff_sign.test(res)
linear_trend.test(res)

irfplot(psi,"IRF")
irfplot(cumsum(psi)+1,"Log IRF")
acf(res)
pacf(res)
```
```{r}
#turning_point.test(res)
#diff_sign.test(res)
#linear_trend.test(res)
#Box.test(res,lag=20,type="L")
adf.test(diff(tsc,lag=4))$p.value
pp.test(diff(tsc,lag=4))$p.value
kpss.test(diff(tsc,lag=4))$p.value  
```

Model 6:
```{r}
w1=pi;
t1=t*t
model5=lm((tsc)~t+t1+cos(w1*t)+t1:sin(w1*t))
summary(model5)

model4<-model5
plot(model4,which=1:2)
sres4=residuals(model4)/sqrt(144.6)
boxplot(sres4)
library(nortest)
ad.test(sres4)
cvm.test(sres4)
lillie.test(sres4)
pearson.test(sres4)
sf.test(sres4)
library(tseries)
res4=residuals(model4)
Box.test(res4)
res4=residuals(model4)
turning_point.test<-function(x)
{
     n<-length(x)
    turns<-0
    for(i in 2:(n-1))
    if((x[i-1]<x[i] && x[i]>x[i+1]) || (x[i-1]>x[i] && x[i]<x[i+1]))
    turns<-turns+1
    z<-(turns-2*(n-2)/3)/sqrt((16*n-29)/90)
    return(list(nturns=turns,z=z,p=2*(1-pnorm(abs(z)))))
     }
turning_point.test(res4)

diff_sign.test<-function(x)
{
n<-length(x)
y<-diff(x)
t<-length(y[y>0])
z<-(t-(n-1)/2)/sqrt((n+1)/12)
return(list(t=t,z=z,p=2*(1-pnorm(abs(z)))))
}
diff_sign.test(res4)

linear_trend.test<-function(x)
{
n<-length(x)
t<-0
for(i in 1:(n-1))
for(j in (i+1):n)
if(x[j]>x[i]) t<-t+1
z<-(t-n*(n-1)/4)/sqrt((n*(n-1)*(2*n+5)/72))
return(list(t=t,z=z,p=2*(1-pnorm(abs(z)))))
}
linear_trend.test(res4)
```

```{r}
model1<-lm(tsc~t+t1+cos(w1*t))
summary(model1)
plot(model1,which=1:2)
sres1<-residuals(model1)/sqrt(494.8)
boxplot(sres1)
plot(t,sres1)
lines(lowess(t,sres1),col=2)
library(nortest)
ad.test(sres1)
cvm.test(sres1)
lillie.test(sres1)
pearson.test(sres1)
sf.test(sres1)
```
```{r}
model6<-tslm(tsc~trend+seasonality)
summary(model6)
plot(model1,which=1:2)
sres1<-residuals(model6)/sqrt(494.8)
boxplot(sres1)
plot(t,sres1)
lines(lowess(t,sres1),col=2)
library(nortest)
ad.test(sres1)
cvm.test(sres1)
lillie.test(sres1)
pearson.test(sres1)
sf.test(sres1)
```


2. Build an appropriate SARIMA model for the same, and iterprete the fitted model interms of its IRF, ACF, PACF and η2n’s. Do the residulas of this SARIMA model satisfy all the required assumptions?                                                                             [8+8+4=20]

SARIMA Model:
```{r}
t=1:66
plot(t,tsc,type="n",main="Community Service",xlab="Quarters",ylab="")
lines(t,tsc)
points(t,tsc,pch=20)
legend("topleft", c("Social (Observed)"), lty=c(1,1), col=c('black','blue'), bty='n', cex=1)
acf(tsc,lag.max=40)
pacf(tsc,lag.max=60)
q1=tsc[seq(1,66,4)]
q2=tsc[seq(2,66,4)]
q3=tsc[seq(3,66,4)]
q4=tsc[seq(4,66,4)]
plot(c(1996,2013),range(q1),type="n",xlab="Year",ylab="Quarterly GDP",main="Quarterwise Plots")
lines(1996:2012,q1)
points(1996:2012,q1)
lines(1996:2012,q2,col=2)
points(1996:2012,q2,col=2)
lines(1996:2011,q3,col=3)
points(1996:2011,q3,col=3)
lines(1996:2011,q4,col=4)
points(1996:2011,q4,col=4)
legend("topleft",c("Q1","Q2","Q3","Q4"),lty=c(1,1),col=1:4,bty='n', cex=1)
```


```{r}
# Unit root test for quarters
pp.test(q1)$p.value
pp.test(q2)$p.value
pp.test(q3)$p.value
pp.test(q4)$p.value
adf.test(q1)$p.value
adf.test(q2)$p.value
adf.test(q3)$p.value
adf.test(q4)$p.value
kpss.test(q1)$p.value
kpss.test(q2)$p.value
kpss.test(q3)$p.value
kpss.test(q4)$p.value
```
```{r}
ag<-tsc
# seasonal difference at lag 4
plot(diff(ag,lag=4),ylab="Seasonal Difference",main="Time Series Plot of Seasonal First Difference")
adf.test(diff(ag,lag=4))$p.value
pp.test(diff(ag,lag=4))$p.value
kpss.test(diff(ag,lag=4))$p.value
acf(diff(ag,lag=4))
pacf(diff(ag,lag=4))
# Ordinary difference in the seasonaaly differenced series
plot(diff(diff(ag,lag=4)),ylab="First Difference (Seasonal+Usual)",main="Time Series Plot of the Differenced Series")
adf.test(diff(diff(ag,lag=4)))$p.value
pp.test(diff(diff(ag,lag=4)))$p.value
kpss.test(diff(diff(ag,lag=4)))$p.value
acf(diff(diff(ag,lag=4)))
pacf(diff(diff(ag,lag=4)))
# test for absence of serial correlation
Box.test(diff(diff(ag,lag=4)),lag=30)
Box.test(diff(diff(ag,lag=4)),lag=30,type="L")
```

```{r}
# SARIMA model selected based on AIC value (least magnitude)
model=arima(ag,order=c(0,1,1),seasonal=list(order=c(1,1,2),period=4))
res_model=residuals(model)
plot(res_model)
qqnorm(res_model)
acf(res_model)
pacf(res_model)
tsdiag(model)
# Normality test for residuals
normtest<-function(x)
{
  library(nortest)
  ad<-ad.test(x)
  s<-shapiro.test(x)
  cvm<-cvm.test(x)
  ll<-lillie.test(x)
  sf<-sf.test(x)
  df<-data.frame(Method=c(s$method,ad$method,cvm$method,ll$method,sf$metho),
                 P.value=c(s$p.value,ad$p.value,cvm$p.value,ll$p.value,sf$p.value))
  df
}
normtest(res_model)
#Portmanteau test for residuals
Box.test(res_model,lag=20)
Box.test(res_model,lag=20,type="L")
```
```{r}
acf(diff(tsc,lag=10))
pacf(diff(tsc,lag=4))
spectrum(diff(tsc,lag=4),spans=c(5,5))
round(1/locator()$x)

acf(diff(diff(tsc,lag=4)))
pacf(diff(diff(tsc,lag=4)))
spectrum(diff(diff(tsc,lag=4)),spans=c(5,5))
#round(1/locator()$x)

```
```{r}
arima(tsc,order=c(1,1,0),seasonal=list(order=c(1,1,0),period=4))
arima(tsc,order=c(1,1,0),seasonal=list(order=c(1,1,0),period=4),fixed=c(NA,0))

arima(tsc,order=c(1,1,0),seasonal=list(order=c(0,1,0),period=4))


arima(tsc,order=c(1,1,0),seasonal=list(order=c(0,1,1),period=4))
arima(tsc,order=c(1,1,0),seasonal=list(order=c(1,1,1),period=4))
arima(tsc,order=c(1,1,0),seasonal=list(order=c(1,1,1),period=4),fixed=c(NA,0,NA))

#arima(tsc,order=c(1,1,0),seasonal=list(order=c(1,0,1),period=4))
arima(tsc,order=c(1,1,0),seasonal=list(order=c(0,0,1),period=4))

```
```{r}
arima(tsc,order=c(0,1,1),seasonal=list(order=c(1,1,0),period=4))
arima(tsc,order=c(0,1,1),seasonal=list(order=c(1,1,0),period=4),fixed=c(NA,0))

arima(tsc,order=c(0,1,1),seasonal=list(order=c(0,1,0),period=4))


arima(tsc,order=c(0,1,1),seasonal=list(order=c(0,1,1),period=4))
arima(tsc,order=c(0,1,1),seasonal=list(order=c(0,1,1),period=4),fixed=c(NA,0))

arima(tsc,order=c(0,1,1),seasonal=list(order=c(1,1,1),period=4))
#arima(tsc,order=c(0,1,1),seasonal=list(order=c(1,1,1),period=4),fixed=c(NA,0,NA))

#arima(tsc,order=c(1,1,0),seasonal=list(order=c(1,0,1),period=4))
arima(tsc,order=c(0,1,1),seasonal=list(order=c(0,0,1),period=4))
```
```{r}
arima(tsc,order=c(0,1,1),seasonal=list(order=c(2,1,0),period=4))
arima(tsc,order=c(0,1,1),seasonal=list(order=c(2,1,0),period=4),fixed=c(NA,0,NA))

arima(tsc,order=c(0,1,1),seasonal=list(order=c(2,1,1),period=4))
arima(tsc,order=c(0,1,1),seasonal=list(order=c(2,1,1),period=4),fixed=c(NA,NA,0,NA))

arima(tsc,order=c(0,1,1),seasonal=list(order=c(0,1,2),period=4))
arima(tsc,order=c(0,1,1),seasonal=list(order=c(0,1,2),period=4),fixed=c(NA,0,NA))


arima(tsc,order=c(0,1,1),seasonal=list(order=c(1,1,2),period=4))
arima(tsc,order=c(0,1,1),seasonal=list(order=c(1,1,2),period=4),fixed=c(NA,NA,NA,0))


```

```{r}
arima(tsc,order=c(1,1,0),seasonal=list(order=c(2,1,0),period=4))$aic
arima(tsc,order=c(1,1,0),seasonal=list(order=c(2,1,1),period=4))$aic
arima(tsc,order=c(1,1,0),seasonal=list(order=c(0,1,2),period=4))$aic
arima(tsc,order=c(1,1,0),seasonal=list(order=c(0,1,2),period=4),fixed=c(NA,0,NA))
arima(tsc,order=c(1,1,0),seasonal=list(order=c(1,1,2),period=4))$aic


############################################

arima(tsc,order=c(0,1,1),seasonal=list(order=c(2,1,0),period=4))$aic
arima(tsc,order=c(0,1,1),seasonal=list(order=c(2,1,1),period=4))$aic
arima(tsc,order=c(0,1,1),seasonal=list(order=c(0,1,2),period=4))$aic
arima(tsc,order=c(0,1,1),seasonal=list(order=c(1,1,2),period=4))$aic

```
```{r}
# Model Testing
aic1<-c(arima(tsc,order=c(0,1,0),seasonal=list(order=c(0,1,0),period=4))$aic)
aic2<-arima(tsc,order=c(0,1,0),seasonal=list(order=c(0,1,1),period=4))$aic
aic3<-arima(tsc,order=c(0,1,0),seasonal=list(order=c(1,1,0),period=4))$aic
aic4<-arima(tsc,order=c(0,1,0),seasonal=list(order=c(1,1,1),period=4))$aic
aic5<-arima(tsc,order=c(0,1,1),seasonal=list(order=c(0,1,0),period=4))$aic
aic6<-arima(tsc,order=c(1,1,0),seasonal=list(order=c(0,1,0),period=4))$aic
aic7<-arima(tsc,order=c(1,1,1),seasonal=list(order=c(0,1,0),period=4))$aic
aic8<-arima(tsc,order=c(0,1,1),seasonal=list(order=c(0,1,1),period=4))$aic
aic9<-arima(tsc,order=c(0,1,1),seasonal=list(order=c(1,1,0),period=4))$aic
aic10<-arima(tsc,order=c(1,1,0),seasonal=list(order=c(1,1,0),period=4))$aic
aic11<-arima(tsc,order=c(1,1,0),seasonal=list(order=c(1,1,1),period=4))$aic
aic12<-arima(tsc,order=c(0,1,1),seasonal=list(order=c(1,1,1),period=4))$aic
aic13<-arima(tsc,order=c(1,1,0),seasonal=list(order=c(0,1,1),period=4))$aic
aic14<-arima(tsc,order=c(1,1,1),seasonal=list(order=c(0,1,1),period=4))$aic
aic15<-arima(tsc,order=c(1,1,1),seasonal=list(order=c(1,1,0),period=4))$aic
aic16<-arima(tsc,order=c(1,1,1),seasonal=list(order=c(1,1,1),period=4))$aic
aic17<-arima(tsc,order=c(0,1,0),seasonal=list(order=c(0,1,2),period=4))$aic
aic18<-arima(tsc,order=c(0,1,0),seasonal=list(order=c(2,1,0),period=4))$aic
aic19<-arima(tsc,order=c(0,1,0),seasonal=list(order=c(2,1,1),period=4))$aic
aic20<-arima(tsc,order=c(0,1,0),seasonal=list(order=c(1,1,2),period=4))$aic
print("-------------------AIC 1:20 ------------------------")
aic1
aic2
aic3
aic4
aic5
aic6
aic7
aic8
aic9
aic10
aic11
aic12
aic13
aic14
aic16
aic17
aic18
aic19
aic20

aic21<-arima(tsc,order=c(0,1,0),seasonal=list(order=c(2,1,2),period=4))$aic
aic22<-arima(tsc,order=c(0,1,1),seasonal=list(order=c(0,1,2),period=4))$aic
aic23<-arima(tsc,order=c(0,1,1),seasonal=list(order=c(2,1,0),period=4))$aic
aic24<-arima(tsc,order=c(0,1,1),seasonal=list(order=c(2,1,1),period=4))$aic
aic25<-arima(tsc,order=c(0,1,1),seasonal=list(order=c(1,1,2),period=4))$aic
aic26<-arima(tsc,order=c(0,1,1),seasonal=list(order=c(2,1,2),period=4))$aic
aic27<-arima(tsc,order=c(0,1,2),seasonal=list(order=c(0,1,1),period=4))$aic
aic28<-arima(tsc,order=c(0,1,2),seasonal=list(order=c(0,1,2),period=4))$aic
aic29<-arima(tsc,order=c(0,1,2),seasonal=list(order=c(1,1,0),period=4))$aic
aic30<-arima(tsc,order=c(0,1,2),seasonal=list(order=c(2,1,0),period=4))$aic
aic31<-arima(tsc,order=c(0,1,2),seasonal=list(order=c(2,1,1),period=4))$aic
aic32<-arima(tsc,order=c(0,1,2),seasonal=list(order=c(1,1,1),period=4))$aic
aic33<-arima(tsc,order=c(0,1,2),seasonal=list(order=c(1,1,2),period=4))$aic
aic34<-arima(tsc,order=c(0,1,2),seasonal=list(order=c(2,1,1),period=4))$aic
aic35<-arima(tsc,order=c(2,1,0),seasonal=list(order=c(2,1,1),period=4))$aic
aic36<-arima(tsc,order=c(2,1,0),seasonal=list(order=c(2,1,0),period=4))$aic
aic37<-arima(tsc,order=c(2,1,0),seasonal=list(order=c(2,1,2),period=4))$aic
aic38<-arima(tsc,order=c(2,1,0),seasonal=list(order=c(0,1,2),period=4))$aic
aic39<-arima(tsc,order=c(2,1,0),seasonal=list(order=c(0,1,1),period=4))$aic
aic40<-arima(tsc,order=c(2,1,0),seasonal=list(order=c(1,1,0),period=4))$aic
print("-------------------AIC 21:40 ------------------------")
aic21
aic22
aic23
aic24
aic25
aic26
aic27
aic28
aic29
aic30
aic31
aic32
aic33
aic34
aic35
aic36
aic37
aic38
aic39
aic40


aic41<-arima(tsc,order=c(2,1,0),seasonal=list(order=c(1,1,1),period=4))$aic
aic42<-arima(tsc,order=c(2,1,1),seasonal=list(order=c(1,1,1),period=4))$aic
aic43<-arima(tsc,order=c(2,1,1),seasonal=list(order=c(0,1,1),period=4))$aic
aic44<-arima(tsc,order=c(2,1,1),seasonal=list(order=c(0,1,0),period=4))$aic
aic45<-arima(tsc,order=c(2,1,1),seasonal=list(order=c(1,1,0),period=4))$aic
aic46<-arima(tsc,order=c(2,1,1),seasonal=list(order=c(1,1,2),period=4))$aic
aic47<-arima(tsc,order=c(2,1,1),seasonal=list(order=c(2,1,2),period=4))$aic
aic48<-arima(tsc,order=c(2,1,1),seasonal=list(order=c(2,1,1),period=4))$aic
aic49<-arima(tsc,order=c(2,1,1),seasonal=list(order=c(2,1,0),period=4))$aic
aic50<-arima(ag,order=c(2,1,1),seasonal=list(order=c(0,1,2),period=4))$aic
aic51<-arima(tsc,order=c(0,1,2),seasonal=list(order=c(0,1,2),period=4))$aic
aic52<-arima(tsc,order=c(1,1,2),seasonal=list(order=c(0,1,2),period=4))$aic
aic53<-arima(tsc,order=c(1,1,2),seasonal=list(order=c(1,1,2),period=4))$aic
aic54<-arima(tsc,order=c(1,1,2),seasonal=list(order=c(1,1,1),period=4))$aic
aic55<-arima(tsc,order=c(1,1,2),seasonal=list(order=c(2,1,1),period=4))$aic
aic56<-arima(tsc,order=c(2,1,2),seasonal=list(order=c(2,1,1),period=4))$aic
aic57<-arima(tsc,order=c(2,1,2),seasonal=list(order=c(0,1,1),period=4))$aic
aic58<-arima(tsc,order=c(2,1,2),seasonal=list(order=c(0,1,2),period=4))$aic

print("-------------------AIC 41:58 ------------------------")
aic41
aic42
aic43
aic44
aic45
aic46
aic47
aic48
aic49
aic50
aic51
aic52
aic53
aic54
aic55
aic56
aic57
aic58



```
Selected Model
```{r}
model<-arima(x = tsc, order = c(1, 1, 0), seasonal = list(order = c(0, 1, 2), period = 4), fixed = c(NA, 0, NA))
summary(model)
res<-residuals(model)
plot(res)
qqnorm(res)
acf(res)
pacf(res)
tsdiag(model)

adf.test(res)
pp.test(res)
normtest(res)
Box.test(res,lag=20,type="L")
BO
```
```{r}
irfplot<-function (irf, s)
 {
 n <- length(irf)
 plot(c(0,n+1), range(c(irf,1)), type = "n",xlab = "Time", ylab = "IRF", main = s)
 lines(c(0,n+1),c(0,0))
 lines(c(0,0),c(0,1))
 for (i in 1:n)
 lines(c(i,i), c(0, irf[i]))
 }
 psi<-ARMAtoMA(ar=c(-0.6363),lag.max=30)
 par(mfrow = c(1, 2))
 #par(oma = c(0, 4, 8, 0))
 irfplot(psi,"IRF")
 irfplot(cumsum(psi)+1,"Log IRF")
```
```{r}
normtest()

```
```{r}
 psi100<-ARMAtoMA( ar=c(-0.6180) ,lag.max=100)
 #psi100
 gamma0<-(1+sum(psi100^2))*4597
 gamma0
 pc2<-ARMAacf( ar=c(-0.6180), lag.max=30,pacf=T)^2
 v<-gamma0
 
 for(i in 2:31) v[i]<-v[i-1]*(1-pc2[i-1])
 ((1-v/gamma0)*100)[-1]
  pc2*100
  cumsum(pc2*100)
 ((gamma0-4597)/gamma0)*100
 
pr<-predict(chosen_model,n.ahead=4)

print("------------------------")

for(i in 2:31) v[i]<-v[i-1]*(1-pc2[i-1])
((1-v/gamma0)*100)[-1]
pc2*100
cumsum(pc2*100)
((gamma0-5963)/gamma0)*100
 
```


3. Use the above two fitted models for finding interval forecasts for 2012-13:Q3, 2012-13:Q4, 2013-14:Q1 and 2013-14:Q2 and check against their realised values. Compare these forecasts and other diagnostics to comment on the aptness of the above two fitted models.        [4+6=10]

```{r}
regression_model<-model5
#regression_model
pr<-predict(regression_model,n.ahead=4)
#pr
f1<-predict(regression_model,data.frame(t=67,t1=67*67),interval="pred",level=0.95)
f2<-predict(regression_model,data.frame(t=68,t1=68*68),interval="pred",level=0.95)
f3<-predict(regression_model,data.frame(t=69,t1=69*69),interval="pred",level=0.95)
f4<-predict(regression_model,data.frame(t=70,t1=70*70),interval="pred",level=0.95)
f1
f2
f3
f4
f<-c(3160.637,3505.228,3356.249,3563.838)
f6<-c(3050.873,3336.266,3314.246,3778.092)
sarima_model<-model
pr<-predict(sarima_model,n.ahead=4)
pr
ts1[67:70]
pr$pred-2*pr$se
pr$pred+2*pr$se
```
```{r}
summary(regression_model)
summary(sarima_model)

```
```{r}
x=1:4

SARIMA_pr<-predict(sarima_model,n.ahead=4)
reg_pr<-predict(sarima_model,n.ahead=4)

a1<-reg_pr$pred-2*reg_pr$se
a2<-reg_pr$pred+2*reg_pr$se

s1<-SARIMA_pr$pred-2*SARIMA_pr$se
s2<-SARIMA_pr$pred+2*SARIMA_pr$se


plot(x,ts1[67:70], type="l", col="blue", main = "Regression Model vs SARIMA model",ylim = c(3160, 4000),xlab="time",ylab="gdp")

points(x,f, col="red", type="b", lty=2)
#points(x,f6, col="yellow", type="b", lty=2)
points(x,SARIMA_pr$pred, col="green", type="b", lty=2)



legend("top", legend=c("observed value", "predicted value:Regression Model","predicted value: SARIMA"),
       col=c("blue", "red","green"), lty=1:2, cex=0.8,
       box.lty=0)


```


Note:

Q1: 




Q2: Differance: D
    p,q : Locator
    Locator: value=1
    Check model different combination (1,1,0)...(1,1,1) and (0,1,0)...(2,1,0)
    Select Model: less AIC
    selected Model: Residual Analysis, IRF, ACF, PACF, eta^2 
    Check: assumptions of sarima model
    
    
    



























